
# Functions for data

Now that we know how to write functions for generating figures, we can use the
same concept and apply to preparing and generating data.

```{r, purl=TRUE}
split_gdp_files <- list.files(path = "data-raw", pattern = "gdp-percapita\\.csv$", full.names = TRUE)
split_gdp_list <- lapply(split_gdp_files, read.csv)
gdp <- do.call("rbind", split_gdp_list)
```

```{r}
gather_gdp_data <- function() {
    split_gdp_files <- list.files(path = "data-raw", pattern = "gdp-percapita\\.csv$", full.names = TRUE)
    split_gdp_list <- lapply(split_gdp_files, read.csv)
    gdp <- do.call("rbind", split_gdp_list)
    gdp
}
```

Just like we created a `make_pdf` function to automatically generate PDF files
from our figures, we can create a `make_csv` function to automatically generate
CSV files from our data sets. This might come handy if you want to send your
intermediate datasets to your collaborators or if you want to inspect more
closely that everything is working as it should.

```{r, purl=TRUE}
make_csv <- function(obj, file, ...,  verbose = TRUE) {
    if (verbose) {
        message("Creating csv file: ", file)
    }
    write.csv(obj, file = file, row.names = FALSE, ...)
}
```

### Your turn

Transform into functions these two pieces of code.

```{r, purl=TRUE}
## Turn this into a function
mean_lifeExp_by_cont <- gdp %>% group_by(continent, year) %>%
  summarize(mean_lifeExp = mean(lifeExp)) %>% as.data.frame
```

```{r, purl=TRUE}
## Turn this into a function
latest_lifeExp <- gdp %>% filter(year == max(gdp$year)) %>%
      group_by(continent) %>%
      summarize(latest_lifeExp = mean(lifeExp)) %>%
      as.data.frame
    latest_lifeExp
```

## Long computations

Caching is available in knitr but it can be pretty fragile. If you can't make a
CSV file because you need to store an R object.

```{r, purl=TRUE}
## If you need to save an R object to avoid the repetition of long computations
make_rds <- function(obj, file, ..., verbose = TRUE) {
    if (verbose) {
        message("Creating rds file: ", file)
    }
    saveRDS(obj, file = file)
}
```

Then in your knitr document, you can do:

`gdp <- readRDS(file="data-output/gdp.rds")`

or maybe even:

```{r, eval=FALSE}
if ( !file.exists("data-output/gdp.rds")) {
    gdp <- gather_gdp_data() ## long computation...
    make_rds(gdp, file="data-output/gdp.rds")
}
gdp <- readRDS(file="data-output/gdp.rds")
```
