## Writing functions to generate figures

Intro to Nile dataset

```{r, purl=FALSE, fig.show='hide'}
plot(Nile)
hist(Nile)
mean(Nile)
```

```{r, purl=TRUE, fig.show='hide'}
par(mfrow=c(1, 2))
plot(Nile)
abline(h = mean(Nile), col = "red", lwd = 3, lty = 2)
yrs <- attr(Nile, "tsp")
fit <- loess(Nile ~ seq(yrs[1], yrs[2], 1))
lines(seq(yrs[1], yrs[2]), fitted(fit), col = "blue", lwd = 2)
hist(Nile)
abline(v = mean(Nile), col = "red",  lwd = 3, lty = 2)
```

If we want to make a PDF file of this figure we could do:

```{r, purl=FALSE, eval=FALSE}
pdf(file="nile_flow.pdf", width=8, height=6)

par(mfrow=c(1, 2))
plot(Nile)
abline(h = mean(Nile), col = "red", lwd = 3, lty = 2)
yrs <- attr(Nile,  "tsp")
fit <- loess(Nile ~ seq(yrs[1], yrs[2], 1))
lines(seq(yrs[1], yrs[2]), fitted(fit), col = "blue", lwd = 2)
hist(Nile)
abline(v = mean(Nile), col = "red",  lwd = 3, lty = 2)

dev.off()
```

It's not very neat, as we end up with 2 variables that are global (`yrs` and
`fit`) that we probably won't need anywhere else.

Let's convert this into a function:

```{r, purl=FALSE}
fig_Nile <- function() {
    par(mfrow=c(1, 2))
    plot(Nile, cex = .7)
    abline(h = mean(Nile), col = "red", lwd = 3, lty = 2)
    yrs <- attr(Nile,  "tsp")
    fit <- loess(Nile ~ seq(yrs[1], yrs[2], 1))
    lines(seq(yrs[1], yrs[2]), fitted(fit), col = "blue", lwd = 2)
    hist(Nile, cex = .7)
    abline(v = mean(Nile), col = "red",  lwd = 3, lty = 2)
}
```

so this part gets a little prettier:

```{r, eval=FALSE}
pdf(file="nile_flow.pdf", width=6.5, height=4)
fig_Nile()
dev.off()
```

If you start making a lot of figures, it would be nice to have to repeat this
first and third lines...

Let's create another function that will automate this process:

```{r, purl=TRUE}
make_pdf <- function(expr,  filename, ..., verbose = TRUE) {
    if (verbose) {
        message("Creating: ", filename)
    }
    pdf(file = filename, ...)
    on.exit(dev.off())
    eval.parent(substitute(expr))
}
```

```{r}
make_pdf(fig_Nile(), "nile_flow.pdf", width = 6.5, height = 4)
```

## Your turn

```{r, purl=TRUE}
## Create your own function that generates a plot and use it with make_pdf.

## If you are looking for some inspiration, the code below compares the change
## in life expectancy for Finland and Japan.
finland <- read.csv(file = "data-raw/Finland-gdp-percapita.csv")
japan <- read.csv(file="data-raw/Japan-gdp-percapita.csv")
plot(lifeExp ~ gdpPercap, data=finland, col="red")
fit_finland <- lm(lifeExp ~ gdpPercap, data=finland)
abline(fit_finland, col="red", lwd=3, lty=2)
points(japan$gdpPercap, japan$lifeExp, pch=2, col="blue")
fit_japan <- lm(lifeExp ~ gdpPercap, data=japan)
abline(fit_japan, col="blue", lwd=3, lty=2)

```
